---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ftolr

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/475787005.svg)](https://zenodo.org/badge/latestdoi/475787005)
<!-- badges: end -->

ftolr is an R package that provides access to the [Fern Tree of Life (FTOL)](https://fernphy.github.io/) and related datasets.

These data were originally compiled using methods described in this paper:

Nitta JH, Schuettpelz E, Ramírez-Barahona S, Iwasaki W. An open and continuously updated fern tree of life https://doi.org/10.3389/fpls.2022.909768

## Installation

You can install ftolr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fernphy/ftolr")
```

## Examples

### Fern Tree of Life (FTOL)

The `ft_tree()` function loads FTOL into the current R session.
There are several options available to control branch lengths, presence or absence of outgroup, and sampling scheme.

```{r tree-example}
library(ftolr)
# Default: consensus tree with branchlengths in units of time
ft_tree()

# Maximum-likelihood (ML) tree with raw branchlengths and outgroup pruned
ft_tree(
  branch_len = "raw",
  consensus = FALSE,
  drop_og = TRUE
) 
```

### DNA sequences

The DNA sequences used to build the tree are also available.
These can be subset by locus (gene) name and can be formatted as a matrix (DNA alignment) or a list.

```{r seq-example}
# Default: aligned Sanger sequences
ft_seqs()

# Unaligned rbcL
ft_seqs(loci = "rbcL", aligned = FALSE)
```

### FTOL data version

You can check the current version of the [FTOL data](https://github.com/fernphy/ftol_data) included in `ftolr` with `ft_data_ver()`:

```{r data-ver}
ft_data_ver()
```

## Citing

If you use this package, please cite it!

Here is an example:

    FTOL working group. (2022). ftolr: Data for the Fern Tree of Life (FTOL) https://doi.org/10.5281/zenodo.6401660

The example DOI above is for the overall package.

Here is the latest DOI, which you should cite if you are using the
latest version of the package:

[![DOI](https://zenodo.org/badge/475787005.svg)](https://zenodo.org/badge/latestdoi/475787005)

Please also cite the paper describing how the data were generated:

    Nitta JH, Schuettpelz E, Ramírez-Barahona S, Iwasaki W. An open and continuously updated fern tree of life
  
## Licenses

- Code: [MIT](LICENSE)
- Data ([ftol.zip](data-raw/ftol.zip)): [CC0](https://creativecommons.org/publicdomain/zero/1.0/)
