---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ftolr

<!-- badges: start -->
<!-- badges: end -->

ftolr is an R package that provides access to the [Fern Tree of Life (FTOL)](https://fernphy.github.io/) and related datasets.

## Installation

You can install the development version of ftolr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("joelnitta/ftolr")
```

## Examples

### Fern Tree of Life (FTOL)

The `ft_tree()` function loads FTOL into the current R session.
There are several options available to control branch lengths, presence or absence of outgroup, and sampling scheme.

```{r tree-example}
library(ftolr)
# Default: maximum-likelihood (ML) tree with branchlengths in units of time
ft_tree()

# ML tree with raw branchlengths and outgroup pruned
ft_tree(
  branch_len = "raw",
  drop_og = TRUE
) 
```

### DNA sequences

The DNA sequences used to build the tree are also available.
These can be subset by locus (gene) name and can be formatted as a matrix (DNA alignment) or a list.

```{r seq-example}
# Default: aligned Sanger sequences
ft_seqs()

# Unaligned rbcL
ft_seqs(loci = "rbcL", aligned = FALSE)
```

